---
title: "Halloween Mini-Proj"
format: pdf
author: "Nickolas Beam"
---

```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
candy = read.csv(candy_file, row.names=1)
head(candy)
```
> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

## Winpercent
The most interesting variables in the dataset is `winpercentt`. For a given candy this value is the percentage of people who prefer this candy over another randomly chosen candy

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
rownames(candy)
```

```{r}
candy["Skittles wildberry",]$winpercent
```

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```


> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

## A useful function from the skimr package

```{r}
library(skimr)
skimr::skim(candy)
```

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent, breaks=20)
```

```{r}
library(ggplot2)
ggplot(candy) + aes(winpercent) + geom_histogram(bins=10, col="red", fill="forest green")
```

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate.inds <- as.logical(candy$chocolate)
choc.win <- candy[chocolate.inds,]$winpercent

fruity.inds <- as.logical(candy$fruity)
fruity.win <- candy[fruity.inds,]$winpercent

mean(choc.win)
mean(fruity.win)
```

## 3. Candy Ranking

First setup some colors for different candy types

```{r}

my_cols=rep("black", nrow(candy))
my_cols
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
my_cols

```

```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent))+
  geom_col(fill = my_cols)
```

```{r}
ggsave("tmp.png")
```

Now, for the first time, using this plot we can answer questions like:
> Q17. What is the worst ranked chocolate candy?


> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```


The labels on this plot are too squished and overlapping. I am going to turn to the ggrepl package and the `geom_text_repel()` function to help avoid overlapping labels.

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

