---
title: "Class08 Lab"
format: pdf
author: "Nickolas Beam"
---

```{r}
fna.data <- "WisconsinCancer.csv"
wisc.df <- read.csv(fna.data, row.names=1)
#wisc.df
```

We will here remove the first column (the expert diagnosis) as we will no tuse this oin our analysis.

```{r}
#We can use -1 here to remove the first column.
wisc.data <- wisc.df[,-1]
diagnosis <- as.factor(wisc.df[,1])
```

> Q1 How many observations are in this dataset?

```{r}
nrow(wisc.data)
```

> Q2 How many of the observations have a malignant diagnosis?

```{r}
table(wisc.df$diagnosis)
```

> Q3 How many variables/features in the data are suffixed with _mean?

```{r}
colnames(wisc.data)
```

Use grep() to search for 
```{r}
matches <- grep("_mean", colnames(wisc.df))
length(matches)
```

# Principal component analysis

```{r}
colMeans(wisc.data)
apply(wisc.data,2,sd)
```

We need to scale with 'scale=TRUE' argument with prcomp()
```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```


```{r}
plot(wisc.pr)
```

Let's make a PC plot (a.k.a "score plot" or "PC1 vs PC2" etc. plot)

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = diagnosis)
```

```{r}
y <- summary(wisc.pr)
attributes(y)
```
How many PCs do I need to capture 80% of the original variance in my dataset?

```{r}
sum(y$importance[3,] <= 0.8)
```



```{r}
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis
library(ggplot2)
ggplot(df) + aes(PC1, PC2, col=diagnosis) + geom_point()
```

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
pve <- pr.var/sum(pr.var)
plot(pve, xlab = "Principal Component",
     ylab = "Proportion of Variance Explained",
     ylim = c(0,1), type = "o")
```

```{r}
barplot(pve, ylab = "Percent of Variance Explained",
        names.arg=paste0("PC", 1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100)
```



## Combine PCA with clustering

I want to cluster in "PC space". 

```{r}
plot(wisc.pr$x[,1],wisc.pr$x[,2], col=diagnosis)
```

```{r}
summary(wisc.pr)
```
The `hclust()` function wants a distance matrix as input...
```{r}
d <- dist(wisc.pr$x[,1:3])
wisc.pr.hclust <- hclust(d, method = "ward.D2")
plot(wisc.pr.hclust)
```
Find my cluster membership vector with the `cutree()` function.

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```


```{r}
table(diagnosis,grps)
```



